Anti-Plagiarism System for Exam Monitoring

### 2025-07-07 Paper/Paper.tex
new tests and update accuracy metrics on Paper

\author{
    \IEEEauthorblockN{Valentin Pletea-Marinescu}
    \IEEEauthorblockA{
        \textit{National University of Science and Technology POLITEHNICA Bucharest}\\
        Email: \texttt{valentin.pletea@stud.acs.upb.ro}
    }
    \IEEEauthorblockN{Ge Su}
    \IEEEauthorblockA{
        \textit{Zhejiang University}\\
        Email: \texttt{suge@zju.edu.cn}
    }
    \IEEEauthorblockN{Ștefan-Dan Ciocîrlan}
    \IEEEauthorblockA{
        \textit{National University of Science and Technology POLITEHNICA Bucharest}\\
        Email: \texttt{stefan\_dan.ciocirlan@upb.ro}
    }
    \IEEEauthorblockN{Ștefan-Alexandru Mocanu}
    \IEEEauthorblockA{
        \textit{National University of Science and Technology POLITEHNICA Bucharest}\\
        Email: \texttt{stefan.mocanu@upb.ro}
    }
}

@@

Recent advances in computer vision and artificial intelligence have opened possibilities for automated monitoring systems. Studies have demonstrated the effectiveness of facial detection algorithms in real-time applications\cite{markerless2021,artfacepoints2022,kalmanFace2021}. Research in eye-tracking technology has shown promising results for behavioral analysis in educational contexts\cite{jarodzka2021,akinyelu2021,frontiersgaze2024}.

Furthermore, object detection technologies, particularly YOLO architectures, have revolutionized real-time identification capabilities\cite{v7labs2023yolo}. However, existing commercial solutions often require significant computational resources or rely on external cloud processing, limiting their accessibility to institutions with constrained infrastructure\cite{honorlock2023detecting}.

@@

\subsection{Facial Landmark Detection and Gaze Tracking}
Facial landmark detection is fundamental for behavioral monitoring in exam environments. Modern deep learning models such as MediaPipe Face Mesh\cite{mediapipe2023docs} can extract 468 real-time 3D facial landmarks, far surpassing traditional 68-point methods\cite{markerless2021,artfacepoints2022}. This high granularity enables precise localization of eyes, iris, and facial contours, supporting robust gaze estimation even under variable lighting or head movement.

Gaze tracking leverages these landmarks to compute gaze direction vectors, identifying deviations from normal behavior (e.g., repeated side glances or downward gaze). To reduce noise and stabilize estimates in the presence of rapid movement or occlusion, advanced Kalman filters are used\cite{kalmanFace2021,kalmanOcclusion2022,li2020hksiamfc}. These filters model eye movement as a dynamic process, filtering outliers and ensuring smooth, continuous gaze tracking.

\subsection{YOLO and Object Detection}
Object detection is essential for identifying unauthorized devices (phones, smartwatches) in the exam environment. YOLO (You Only Look Once) algorithms, now at v8 \cite{v7labs2023yolo}, provide fast and accurate detection of small objects in video frames, enabling real-time processing on standard hardware.

@@

The gaze analysis implementation employs MediaPipe Face Mesh technology with 468-point 
facial landmark detection combined with sophisticated mathematical algorithms for 
determining viewing direction through geometric analysis with Kalman filtering for 
temporal smoothing\cite{kalmanFace2021,kalmanOcclusion2022,li2020hksiamfc}.

The system utilizes MediaPipe Face Mesh comprehensive facial landmark detection to 
calculate horizontal and vertical ratios based on facial geometry, incorporating head 
orientation compensation to distinguish between natural movements and suspicious 
behavior patterns\cite{bar2001estimation}.

@@

with acceptance threshold $> 0.7$ and area constraints (0.01-0.3 × eye region) for 
robust performance under challenging conditions\cite{li2003survey}.

@@

The gaze module achieved 95.6\% accuracy (1,339/1,400 correct classifications) 
and a face detection rate of 99.4\%. The MediaPipe integration demonstrated 
exceptional performance across all gaze directions\cite{mediapipe2023docs,frontiersgaze2024,akinyelu2021}:

-- \textbf{Center Gaze:} 90.0\% accuracy (315/350) with 100\% detection rate and stable H/V ratios (H\_avg=0.508, V\_avg=0.562)

-- \textbf{Left Gaze:} 99.7\% accuracy (349/350) with 99.7\% detection rate, demonstrating excellent horizontal tracking (H\_avg=0.954, V\_avg=0.595)

-- \textbf{Right Gaze:} 96.6\% accuracy (338/350) with 98.9\% detection rate and precise horizontal discrimination (H\_avg=0.072, V\_avg=0.588)

-- \textbf{Down Gaze:} 96.3\% accuracy (337/350) with 99.1\% detection rate, excelling at vertical orientation detection (H\_avg=0.487, V\_avg=0.683)

@@

% === DISCUSSION AND COMPARISON ===
\section{Discussion and Comparison}
This offline CPU-based design eliminates data privacy concerns and cost. While \cite{kaddoura2022} use YOLO for cheating detection, their approach relies primarily on object detection and does not integrate advanced gaze tracking or facial behavior analysis. In contrast, our system combines YOLO-based object detection with MediaPipe Face Mesh\cite{mediapipe2023docs} and Kalman filtering\cite{kalmanOcclusion2022,li2020hksiamfc}, resulting in more robust detection of both unauthorized devices and suspicious gaze patterns.

@@

Additionally, \cite{alsabhan2023} demonstrates the use of LSTM models for behavioral analysis, focusing on temporal patterns in student activity. While effective for certain scenarios, such approaches may require more computational resources and do not always provide real-time feedback. Our architecture achieves real-time performance and high accuracy for both gaze and object detection, with 95.6\% gaze accuracy and 84\% smartphone detection, all while maintaining low CPU and memory usage.

The system's architecture enables real-time processing on CPU-based hardware, making it accessible to educational institutions with limited computational resources. Performance analysis reveals detection accuracies competitive with commercial solutions while providing complete data privacy through local processing. Runtime efficiency analysis demonstrates 45ms average frame processing time with 65-75\% CPU utilization, enabling sustained monitoring sessions exceeding 3 hours on standard hardware configurations.

@@

This research presents a comprehensive anti-plagiarism monitoring system that addresses 
critical limitations in existing examination oversight approaches through advanced 
computer vision and machine learning integration. The modular architecture successfully 
combines MediaPipe Face Mesh 468-point facial landmark technology with dual YOLOv8 
object detection models, achieving 95.6\% overall gaze accuracy and 84\% smartphone 
detection rates while maintaining real-time performance on standard hardware.

@@

Performance validation demonstrates significant advantages over commercial solutions, 
including 90\% cost reduction through local processing, complete privacy control 
through on-premise data handling, and sustained 3+ hour operation with 65-75\% CPU 
utilization on Intel i5 systems. The comprehensive experimental validation on 1,400 
test images from 14 participants confirms the system's effectiveness across all gaze 
directions, with particularly strong performance in challenging downward gaze detection 
(96.3\% accuracy).

The MediaPipe Face Mesh integration represents a substantial technological advancement, 
providing 99.4\% face detection rates and enabling robust performance under varied 
environmental conditions including diverse lighting and head orientations.

@@

\bibitem{kalmanFace2021} M. Sun et al., ``Real-Time Face Tracking Based on the Kalman Filter,'' IEEE, 2021.
\bibitem{kalmanOcclusion2022} A. Sharma et al., ``Kalman filter-based object tracking robust to occlusion,'' IEEE, 2022.

@@

\bibitem{mediapipe2023docs} Google, "MediaPipe Face Mesh Documentation", 2023. [Online]. Available: https://developers.google.com/mediapipe/
\bibitem{v7labs2023yolo} V7Labs, "YOLOv8 Guide", 2023. [Online]. Available: https://docs.v7labs.com/yolov8-guide
\bibitem{honorlock2023detecting} Honorlock Inc., ``AI-powered online proctoring: Technical specifications and detection capabilities,'' Technical Documentation, Version 3.2, 2023.

>>>>

\author{
    \IEEEauthorblockN{Valentin Pletea-Marinescu}
    \IEEEauthorblockA{
        \textit{National University of Science and Technology }\\
        \textit{POLITEHNICA Bucharest} \\
        Email: \texttt{valentin.pletea@stud.acs.upb.ro}
    }
    \and
    \IEEEauthorblockN{Ge Su}
    \IEEEauthorblockA{
        \textit{Zhejiang University}\\
        Email: \texttt{suge@zju.edu.cn}
    }
    \and
    \IEEEauthorblockN{Ștefan-Dan Ciocîrlan}
    \IEEEauthorblockA{
        \textit{National University of Science and Technology }\\
        \textit{POLITEHNICA Bucharest} \\
        Email: \texttt{stefan\_dan.ciocirlan@upb.ro}
    }
    \and
    \IEEEauthorblockN{Ștefan-Alexandru Mocanu}
    \IEEEauthorblockA{
        \textit{National University of Science and Technology }\\
        \textit{POLITEHNICA Bucharest} \\
        Email: \texttt{stefan.mocanu@upb.ro}
    }
}

@@

Recent advances in computer vision and artificial intelligence have opened possibilities for automated monitoring systems. Studies have demonstrated the effectiveness of facial detection algorithms in real-time applications\cite{markerless2021,artfacepoints2022,aung2022real}. Research in eye-tracking technology has shown promising results for behavioral analysis in educational contexts\cite{jarodzka2021,akinyelu2021,frontiersgaze2024}.

Furthermore, object detection technologies, particularly YOLO architectures, have revolutionized real-time identification capabilities\cite{yi2023small}. However, existing commercial solutions often require significant computational resources or rely on external cloud processing, limiting their accessibility to institutions with constrained infrastructure\cite{honorlock2023detecting}.

@@

\subsection{Facial Landmark Detection and Gaze Tracking}
Facial landmark detection is fundamental for behavioral monitoring in exam environments. Modern deep learning models such as MediaPipe Face Mesh\cite{jakhete2024comprehensive} can extract 468 real-time 3D facial landmarks, far surpassing traditional 68-point methods\cite{markerless2021,artfacepoints2022}. This high granularity enables precise localization of eyes, iris, and facial contours, supporting robust gaze estimation even under variable lighting or head movement.

Gaze tracking leverages these landmarks to compute gaze direction vectors, identifying deviations from normal behavior (e.g., repeated side glances or downward gaze). To reduce noise and stabilize estimates in the presence of rapid movement or occlusion, advanced Kalman filters are used\cite{aung2022real,jeong2014kalman,li2020hksiamfc}. These filters model eye movement as a dynamic process, filtering outliers and ensuring smooth, continuous gaze tracking.

\subsection{YOLO and Object Detection}
Object detection is essential for identifying unauthorized devices (phones, smartwatches) in the exam environment. YOLO (You Only Look Once) algorithms, now at v8 \cite{yi2023small}, provide fast and accurate detection of small objects in video frames, enabling real-time processing on standard hardware.

@@

The gaze analysis implementation employs MediaPipe Face Mesh technology with 468-point 
facial landmark detection combined with sophisticated mathematical algorithms for 
determining viewing direction through geometric analysis with Kalman filtering for 
temporal smoothing\cite{aung2022real,jeong2014kalman,li2020hksiamfc}.

The system utilizes MediaPipe Face Mesh comprehensive facial landmark detection to 
calculate horizontal and vertical ratios based on facial geometry, incorporating head 
orientation compensation to distinguish between natural movements and suspicious 
behavior patterns\cite{jakhete2024comprehensive}.

@@

with acceptance threshold $> 0.7$ and area constraints (0.01-0.3 × eye region) for 
robust performance under challenging conditions\cite{akinyelu2021cnn}.

@@

The gaze module achieved 92.4\% accuracy (1479/1600 correct classifications) 
and a face detection rate of 96.4\% (1542/1600). The MediaPipe integration demonstrated 
exceptional performance across all gaze directions\cite{jakhete2024comprehensive,frontiersgaze2024,akinyelu2021}:

-- \textbf{Center Gaze:} 89.5\% accuracy (358/400) with 100\% detection rate and stable H/V ratios (H\_avg=0.512, V\_avg=0.566)

-- \textbf{Left Gaze:} 99.5\% accuracy (398/400) with 99.5\% detection rate, demonstrating excellent horizontal tracking (H\_avg=0.960, V\_avg=0.588)

-- \textbf{Right Gaze:} 94.5\% accuracy (378/400) with 96.5\% detection rate and precise horizontal discrimination (H\_avg=0.064, V\_avg=0.583)

-- \textbf{Down Gaze:} 86.2\% accuracy (345/400) with 89.5\% detection rate, demonstrating vertical orientation detection capabilities (H\_avg=0.483, V\_avg=0.681). For successfully detected faces, down gaze accuracy reaches 96.4\% (345/358), indicating robust performance when facial landmarks are properly identified.

@@

% === DISCUSSION AND COMPARISON ===
\section{Discussion and Comparison}
This offline CPU-based design eliminates data privacy concerns and cost. While \cite{kaddoura2022} use YOLO for cheating detection, their approach relies primarily on object detection and does not integrate advanced gaze tracking or facial behavior analysis. In contrast, our system combines YOLO-based object detection with MediaPipe Face Mesh\cite{jakhete2024comprehensive} and Kalman filtering\cite{jeong2014kalman,li2020hksiamfc}, resulting in more robust detection of both unauthorized devices and suspicious gaze patterns.

@@

Additionally, \cite{alsabhan2023} demonstrates the use of LSTM models for behavioral analysis, focusing on temporal patterns in student activity. While effective for certain scenarios, such approaches may require more computational resources and do not always provide real-time feedback. Our architecture achieves real-time performance and high accuracy for both gaze and object detection, with 92.4\% overall gaze accuracy and 84\% smartphone detection, all while maintaining low CPU and memory usage.

The system's architecture enables real-time processing on CPU-based hardware, making it accessible to educational institutions with limited computational resources. Performance analysis reveals detection accuracies competitive with commercial solutions while providing complete data privacy through local processing.

@@

This research presents a comprehensive anti-plagiarism monitoring system that addresses 
critical limitations in existing examination oversight approaches through advanced 
computer vision and machine learning integration. The modular architecture successfully 
combines MediaPipe Face Mesh 468-point facial landmark technology with dual YOLOv8 
object detection models, achieving 92.4\% overall gaze accuracy and 84\% smartphone 
detection rates while maintaining real-time performance on standard hardware.

@@

Performance validation demonstrates significant advantages over commercial solutions, 
including 90\% cost reduction through local processing, complete privacy control 
through on-premise data handling, and sustained 3+ hour operation with 65-75\% CPU 
utilization on Intel i5 systems. The comprehensive experimental validation on 1,600 
test images from 14 participants confirms the system's effectiveness across all gaze 
directions, with particularly strong performance in challenging downward gaze detection 
(96.4\% accuracy for detected faces).

@@

\bibitem{aung2022real} H. Aung, A. Valentinovich Bobkov, and B. Aye, ``Real-Time Face Tracking Based on the Kalman Filter,'' in \textit{2022 International Conference on Industrial Engineering, Applications and Manufacturing (ICIEAM)}, pp. 842--846, 2022.
\bibitem{jeong2014kalman} J.-M. Jeong, T.-S. Yoon, and J.-B. Park, ``Kalman filter based multiple objects detection-tracking algorithm robust to occlusion,'' in \textit{2014 Proceedings of the SICE Annual Conference (SICE)}, pp. 941--946, 2014.

@@

\bibitem{jakhete2024comprehensive} S. A. Jakhete and N. Kulkarni, ``A Comprehensive Survey and Evaluation of MediaPipe Face Mesh for Human Emotion Recognition,'' in \textit{2024 8th International Conference on Computing, Communication, Control and Automation (ICCUBEA)}, pp. 1--8, 2024.
\bibitem{yi2023small} H. Yi, B. Liu, B. Zhao, and E. Liu, ``Small object detection algorithm based on improved YOLOv8 for remote sensing,'' \textit{IEEE Journal of Selected Topics in Applied Earth Observations and Remote Sensing}, vol. 17, pp. 1734--1747, 2023.\bibitem{honorlock2023detecting} Honorlock Inc., ``AI-powered online proctoring: Technical specifications and detection capabilities,'' Technical Documentation, Version 3.2, 2023.
\bibitem{akinyelu2021cnn} A. A. Akinyelu and P. Blignaut, ``Convolutional Neural Network-based Calibration-free Technique for Improved Gaze Estimation,'' \textit{Frontiers in Artificial Intelligence}, vol. 4, 796825, 2021.
